---
output: github_document
editor_options: 
  chunk_output_type: console
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  dev = "png", dpi = 300,
  fig.height = 3.5, 
  cache = FALSE
)
```

```{r, include = FALSE}
set.seed(1)
```

# biroot

**Note: this package is currently under construction. You should not use it.**
<!-- extends [**ggplot2**](https://github.com/tidyverse/ggplot2) providing -->
<!-- functions to visualize the decision spaces of classification and clustering methods. -->
<!-- **ggclassify** currently includes the functions `geom_classify()` and `geom_classify_boundary()` for plotting the classification regions through filling in the regions, or outlining the resulting -->
<!-- boundaries respectively. -->

## Basic usage

**biroot** supplies tools for finding the one dimensional root of a function of two variables over a fixed rectangular spatial extent. For example, if $f(x,y) = x^{2} + y^{2} - 1$ is the function of interest and $\mathcal{D} = [-2,2] \times [-2,2] \subset \mathbb{R}^{2}$ is the spatial extent, the goal of these tools would be to find that the set $\{(x,y) \in \mathcal{D}: f(x,y) = 0\}$ is the unit circle and provide a data frame parameterizing points on it. 

A few notes are worth mentioning.

1. As is typical in R, the function $f(x,y)$ is not implemented as `f(x,y)`{.R} but rather `f(v)`{.R}, where `v = c(x,y)`{.R}.
2. Points $(x,y)$ such that $f(x,y) = 0$ are referred to as *solutions of $f$*, and the collection of such points is the level set of $f$ at $0$. For an arbitrary $f$, the level set of $f$ at $0$ can be quite complicated: it can have isolated solutions ("0d" solutions, e.g. $f(x,y) = x^{2} + y^{2}$), curves of solutions ("1d" solutions, e.g. $f(x,y) = x^{2} + y^{2} - 1$), spatial extents of solutions ("2d" solutions, e.g. $f(x,y) = 1[x^{2} + y^{2} \leq 1]$), or no solutions (e.g. $f(x,y) = x^{2} + y^{2} + 1$). **biroot** is only defined with 1d solutions in mind.
3. 1d solution sets can be complicated. They can be disconnected (e.g. $f(x,y) = ((x-2)^{2} + y^{2} - 1)((x+2)^{2} + y^{2} - 1)$), unbounded (e.g. $f(x,y) = y - x$), self intersecting (e.g. $f(x,y) = (y-x)(y+x)$), or exhibit other cusp-like behavior (e.g. $f(x,y) = y^{2} - x^{3}$).

The basic algorithm used in biroot is [the quadtree algorithm](https://en.wikipedia.org/wiki/Quadtree), of which there are many variations depending on use casee but here can be seen to be an adaptive grid search, similar to a 2d analogue of the [bisection method](https://en.wikipedia.org/wiki/Bisection_method). It has two core functions:

1. `biroot()`{.R}, which accepts a function `f`{.R} and spatial extents `xlim`{.R} and `ylim`{.R} (numeric vectors of length 2). This function is called `biroot(f, xlim, ylim)`{.R} and returns a data frame of places (`x`{.R} and `y`{.R}) where `f`{.R} was evaluated in the Cartesian product of `xlim`{.R} and `ylim`{.R}, the value of the function at those points (`value`{.R}), along with other information corresponding to the algorithm (`depth`{.R}, or which scale, and `id`{.R}, which rectangle at that scale).
2. `biroot_lines()`{.R} and `biroot_bands()`{.R}, made by analogy to `isolines()`{.R} and `isobands()`{.R} from [the **isoband** package](https://isoband.r-lib.org) and having a similar data structure.

Here is an example of the basic use of these functions.

```{r, fig.height=6, fig.width=6, fig.dpi=300}
library("biroot")

ftest <- function(v) with(v, x^2 + y^2 - 1)
df <- biroot(ftest, xlim = c(-2,2), ylim = c(-2,2)) 
str(df)

color_function <- function(x) {
  # f <- colorRampPalette(c("firebrick1", "gray85", "steelblue1"))(10) |> 
  f <- colorRampPalette(c("gray85", "firebrick1", "gray85"))(10) |> 
    colorRamp()
  
  threshold_to_interval <- function(x, l, u) l*(x <= l) + x*(l < x & x < u) + u*(x >= u)
  
  rescale <- function(x) (x - min(x)) / (max(x) - min(x))
  
  x |> 
    threshold_to_interval(l = -1, u = 1) |> 
    rescale() |> 
    f() |> 
    apply(1, \(v) rgb(v[1], v[2], v[3], maxColorValue = 255))
}

plot(x, y, bg = color_function(value), pch = 21, col = "gray25") |> 
  with(df, expr = _)
```

The boundary argument can also be used to get an aproximation of the root contour
using marching squares. Here's an example of how the aproximation improves as
a larger depth is used.

```{r, fig.height=6, fig.width=6, fig.dpi=300}
library("ggplot2")
theme_set(theme_minimal())
theme_update(panel.grid.minor = element_blank())
library("patchwork")

heartf <- function(df) with(df,x^6 + 3*x^4*y^2 - 3*x^4 + 3*x^2*y^4 - 6*x^2*y^2 + 3*x^2 + y^6 - 3*y^4 + 3*y^2 - 1 - x^2*y^3)

biroot_lines(f = heartf, xlim = c(-1.5,1.5),
       ylim = c(-1,1.5), max_depth = 7) |> 
  ggplot(aes(x,y,group = id))+
  geom_line()+
  ggtitle("depth = 7")+
    theme(axis.text.x = element_blank(), 
    axis.ticks.x = element_blank(),
    axis.title.x = element_blank())+
  coord_equal(xlim = c(-1.25,1.25), ylim = c(-1.15,1.35), expand = FALSE)+
biroot_lines(f = heartf, xlim = c(-1.5,1.5),
         ylim = c(-1,1.5), max_depth = 8) |> 
  ggplot(aes(x,y,group = id))+
  geom_line()+
  ggtitle("depth = 8")+
  theme(axis.text.y = element_blank(), 
    axis.ticks.y = element_blank(),
    axis.title.y = element_blank(),
    axis.text.x = element_blank(), 
    axis.ticks.x = element_blank(),
    axis.title.x = element_blank())+
  coord_equal(xlim = c(-1.25,1.25), ylim = c(-1.15,1.35), expand = FALSE)+
biroot_lines(f = heartf, xlim = c(-1.5,1.5),
         ylim = c(-1,1.5), max_depth = 9) |> 
  ggplot(aes(x,y,group = id))+
  geom_line()+
  ggtitle("depth = 9")+
  coord_equal(xlim = c(-1.25,1.25), ylim = c(-1.15,1.35), expand = FALSE)+
biroot_lines(f = heartf, xlim = c(-1.5,1.5),
         ylim = c(-1,1.5), max_depth = 10) |> 
  ggplot(aes(x,y,group = id))+
  geom_line()+
  ggtitle("depth = 10")+
  theme(axis.text.y = element_blank(), 
    axis.ticks.y = element_blank(),
    axis.title.y = element_blank())+
  coord_equal(xlim = c(-1.25,1.25), ylim = c(-1.15,1.35), expand = FALSE)+
  plot_layout(ncol = 2)
```



## Installation

You can install the latest development version of **ggclassify** from [GitHub](https://github.com/) with:

``` r
if (!requireNamespace("remotes")) install.packages("remotes")
remotes::install_github("nathaniel-morgan/biroot")
```



